---
title: "Test exam numerical question 1"
author: "Stan Koobs"
date: "1-7-2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages(c("tidyverse","ggplot2","MASS","mvtnorm","readxl", "boot","doParallel"))
```

## R Markdown

We first make a plot of these functions. We use a theme designed for this exam :).

```{r }

Numericalmethodstheme <- function() {
  theme(panel.background = element_rect(fill = "grey96"),
        axis.line = element_blank(),
        panel.border = element_rect(colour = "gray82", fill = NA,
                                    size = 1.5),
        panel.grid.major = element_line(colour = "grey70", 
                                        linetype = "dashed"),
        panel.grid.minor = element_line(colour = "grey70", 
                                        linetype = "dashed"),
        axis.text.x = element_text(size= 13),
        axis.text.y = element_text(size= 13),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        axis.ticks = element_line(size = 1.5),
        axis.ticks.length = unit(1.5, "mm"),
        legend.position = "none"
  )
}

xvec = seq(0, 30, by = 0.01)

curvesplot = ggplot() +              
  geom_line(aes(x = xvec, y = dchisq(xvec, df = 4), 
                color = "Chi-square")) +
  geom_line(aes(x = xvec, y = df(xvec, df1 = 3, df2 = 4), 
                color = "F")) +
  scale_color_manual(name = "Legend", 
                     values = c("Chi-square" = "#FF0000", 
                                "F" = "#0000FF")) +
  labs(title = "Chi-square and F distribution", 
       y = "y", x = "x")
  

curvesplot


```
We observe an intersection around zero, around 2 and maybe around the end. Let us use a root-finding algorithm to find these. We first define the bisection method.


```{r}


bisection <- function(ftn, x.l, x.r, tol = 1e-6, max.iter = 100) {

  # check inputs
  if (x.l >= x.r) {
    cat("error: x.l >= x.r \n")
    return(NULL)
  }
  f.l <- ftn(x.l)
  f.r <- ftn(x.r)
  if (f.l == 0) {
    return(x.l)
  } else if (f.r == 0) {
    return(x.r)
  } else if (f.l * f.r > 0) {
    cat("error: ftn(x.l) * ftn(x.r) > 0 \n")
    return(NULL)
  }
  # successively refine x.l and x.r
  n <- 0
  while ((x.r - x.l) > tol && n < max.iter) {
    x.m <- (x.l + x.r) / 2
    f.m <- ftn(x.m)
    if (f.m == 0) {
      return(x.m)
    } else if (f.l * f.m < 0) {
      x.r <- x.m
      f.r <- f.m
    } else {
      x.l <- x.m
      f.l <- f.m
    }
    n <- n + 1
    cat("at iteration", n, "the root lies between", x.l, "and", x.r, "\n")
  }
  # return (approximate) root
  return((x.l + x.r) / 2)
}


```


Now we define difference of the two functions. 
```{r}

rootfunc = function(x) {
  dchisq(x, df = 4) - df(x, df1 = 3, df2 = 4)
}

```

We will first search for the root around 2. Therefore we use the interval [1,3] in the bisection method.

```{r}

bisection(rootfunc, 1, 3)

```
Furthermore, also note that the another intersection can be found at the zero point.


```{r}

rootfunc(0)

```

We will make another plot to check what happens between 10 and 30 since we cannot see that from the previous.


```{r}

xvec = seq(0, 30, by = 0.01)

ggplot() +
  geom_line(aes(x = xvec, y = rootfunc(xvec))) +
  geom_hline(yintercept = 0) +
  xlim(10, 30) +
  ylim(-0.005, 0.01)

```
We see another root around 16. We now use the bisection method another time.

```{r}

bisection(rootfunc, 10, 20)

```
We find another root at 16.32375.

We can do one more check to the right of this interval.


```{r}
bisection(rootfunc, 16.33, 10000000)
```

Now we indeed get an error that the line does not cross the x-axis anymore. The probability mass of both functions is so extremely small there that we can be sure that we can not find any further intersections.

We conclude that the intersection of f(x) and g(x) can be found at x = 0, x = 1.6835 and x = 16.32375.




